Thesis outline

### 2017-06-16 KellyDouglass/Thesis/Thesis_sections/Paper3.tex
Updated Paper 3

We use the Direct $T_e$ method described in \cite{Izotov06} to estimate the 
gas-phase abundances of oxygen and nitrogen in our sample of dwarf galaxies, 
because this method is often regarded as the most accurate estimate of element 
abundances.  It can be difficult to use due to the nature of the [\ion{O}{3}] 
$\lambda$4363 auroral line \citep[for a more detailed discussion, see][]
{Douglass17a}.  Since the strength of [\ion{O}{3}] $\lambda$4363 is inversely 
proportional to the metallicity of a galaxy, it is best suited for low-redshift, 
low-metallicity galaxies.  At metallicities 
$12 + \log (\text{O}/\text{H}) \gtrsim 8.5$, [\ion{O}{3}] $\lambda$4363 becomes 
too weak to detect in the SDSS spectra.  With the mass-metallicity (MZ) relation 
by \cite{Tremonti04}, this metallicity limit corresponds to a maximum stellar 
mass of
\begin{align*}
    12 + \log \left( \frac{\text{O}}{\text{H}} \right) &= -1.492 + 1.847(\log M_*) - 0.08026(\log M_*)^2\\
    8.5 &= \\
    \log \left( \frac{M_*}{M_\odot} \right) &= 8.696
\end{align*}
The MZ relation by \cite{Andrews13} estimates that this maximum metallicity 
translates to a maximum stellar mass of
\begin{align*}
    12 + \log \left( \frac{\text{O}}{\text{H}} \right) &= 12 + \log \left( \frac{\text{O}}{\text{H}} \right)_{asm} - \log \left( 1 + \left( \frac{M_{TO}}{M_*} \right)^\gamma \right)\\
    8.5 &= 8.798 - \log \left( 1 + \left( \frac{10^{8.901}}{M_*} \right)^{0.640} \right)\\
    M_* &= 8.138\times 10^8 M_\odot\\
    \log \left( \frac{M_*}{M_\odot} \right) &= 8.911
\end{align*}
These upper limits on the galaxy stellar masses correspond to the maximum mass 
of a dwarf galaxy in SDSS ($\log (M_*/M_\odot) \approx 9$).  Therefore, we can 
expect to estimate the chemical abundances of only dwarf galaxies in SDSS DR7 
with the Direct $T_e$ method.  Higher resolution spectra would be necessary to 
probe higher mass, higher metallicity galaxies.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{Images/Paper3/Zlow_v_Zreal_1sig_I06_BPTclass_fits}
    \caption{$12 + \log{\text{O}^{++}/\text{H}}$ versus 
    $12 + \log{\text{O}/\text{H}}$ for those galaxies with $M_r > -20$ as 
    calculated with the direct method; the galaxies are colored by their 
    classification in the BPT diagrams (from \cite{Brinchmann04}).  Two linear 
    models have been fit to three groups in the sample: star-forming and 
    AGN/composite galaxies; the best fit parameters can be found in Table 
    \ref{tab:ab}.}
    \label{fig:lowVreal}
\end{figure}

We derive an ICF for O$^+$ to overcome some limitations on the observations of 
dwarf galaxies in SDSS to bolster our sample size.  Because we are studying 
dwarf galaxies, the [\ion{O}{2}] $\lambda 3727$ doublet necessary for estimating 
the abundance of O$^+$ is only available for objects within the redshift range 
$0.02 < z < 0.03$ \citep[see Sec. \ref{sec:SDSS_limits} and][for more 
details]{Douglass17a}.  To be able to study SDSS galaxies at redshifts less than 
0.02, we need to find an alternate way to estimate the abundance of O$^+$.  
Ordinarily, we would use an ICF to replace the missing ionization stage, as 
outlined in Sec. \ref{sec:DirectTe} for the nitrogen abundance.  Because the 
[\ion{O}{2}] $\lambda 3727$ emission line is normally available for analysis, 
there are no approximations available for the ICF for the amount of 
singly-ionized oxygen.  To construct an ICF appropriate for our dwarf galaxy 
sample, we compare $12 + \log{\text{O/H}}$ to 
$12 + \log{\text{O}^{++}/\text{H}}$ as calculated by the Direct $T_e$ method for 
all galaxies in SDSS with $M_r > -20$.  Recognizing that the relative amounts of 
singly- and doubly-ionized oxygen will depend on the hardness of a galaxy's 
spectrum, we use a Baldwin-Phillips-Terlevich (BPT) diagram \citep{Baldwin81} to 
classify each galaxy as star-forming, AGN, or composite (containing 
characteristics of both a star-forming region and an AGN).  Noting a strong 
separation in the relationship between O$^{++}$/H and O/H with respect to the 
galaxies' BPT classifications by \cite{Brinchmann04}, we fit two linear models 
to the sample, seen in Fig. \ref{fig:lowVreal} and defined as 
\begin{equation}\label{eq:fit}
    12 + \log{\text{O}/\text{H}} = a(12 + \log{\text{O}^{++}/\text{H}}) + b
\end{equation}
The values for $a$ and $b$ for the two classes are listed in Table \ref{tab:ab}.

\floattable
\begin{deluxetable}{L|C|C|C}
    \tablewidth{0pt}
    \tablecaption{Coefficients of oxygen abundance fits\label{tab:ab}}
    \tablehead{\colhead{} & \colhead{SF (dwarf)} & \colhead{SF ($M_r > -20$)} & \colhead{AGN \& Composite}}
    \startdata
        a & 0.84 \pm 0.013 & 0.824 \pm 0.0010 & 0.764 \pm 0.0035\\
        b & 1.61 \pm 0.094 & 1.787 \pm 0.0065 & 2.05 \pm 0.023\\
    \enddata
    \tablecomments{Coefficients for the linear trends (Eqn. \ref{eq:fit}) fit to 
    the SF galaxies (dwarf and those with $M_r > -20$) and AGN and composite 
    galaxies.  With these trends, we can now calculate the total oxygen 
    abundance for a galaxy with knowing only the O$^{++}$/H abundance.}
\end{deluxetable}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%    DATA
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{SDSS data and galaxy selection}

@@

%-------------------------------------------------------------------------------
\subsection{Spectroscopic selection}\label{sec:SDSS_limits}

@@

%-------------------------------------------------------------------------------
\subsection{Void classification}
% Re-write to avoid plagarizing myself :P

Void galaxies are identified using the void catalog compiled by \cite{Pan12}, 
which was built based on the galaxies in SDSS DR7 catalog.  Starting with 
galaxies with absolute magnitudes $M_r < -20$, the VoidFinder algorithm of 
\cite{Hoyle02} \citep[based on the algorithm described by][]{ElAd97} removes all 
isolated galaxies (defined as having the third nearest neighbor more than 7 
$h^{-1}$ Mpc away).  After applying a grid to the remaining galaxies, spheres 
are grown from all empty grid cells (cells containing no galaxies).  A sphere 
reaches its maximum size when it encounters four galaxies on its surface.  To be 
classified as a void (or part of one), a sphere must have a minimum 10 Mpc 
radius.  If two spheres overlap by more than 10\%, they are considered part of 
the same void.  See \cite{Hoyle02} for a more detailed description of the 
VoidFinder algorithm.  Those galaxies that fall within these void spheres are 
classified as voids.  Those galaxies that lie outside the spheres are classified 
as wall galaxies.  Because we cannot identify any voids within 10 Mpc of the 
edge of the survey, we do not include the galaxies that fall within this region 
in either the void or wall sample (throughout this paper, these galaxies are 
labeled as ``Uncertain'').

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%    ANALYSIS & RESULTS
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Abundance analysis and results}

@@

%Our primary objective is to perform a relative measurement of metallicity of 
%dwarf galaxies to discern how the large-scale environment affects their chemical 
%evolution.  As discussed in Section \ref{sec:Theory}, the strength of and 
%ability to observe different spectral lines between various surveys and 
%observations require multiple methods to be developed for metallicity 
%calculations.  In this paper, we use only the Direct $T_e$ method, because no 
%other method has yet been calibrated using dwarf galaxies.  The results from the 
%various methods are not directly comparable; while they all return metallicities 
%within the same range, the same galaxy can have very different metallicity 
%values depending on which method is used.  Conversions between methods have been 
%developed \citep[see][]{Kewley08}, but it is not clear that these conversions 
%would be accurate for dwarf galaxies.  Unfortunately, there are not enough 
%galaxies available in our sample to calibrate these other methods for dwarf 
%galaxies.

All line ratios listed are ratios of the emission-line fluxes.  Galaxies with 
low metallicities have $Z = 12 + \log (\text{O}/\text{H}) < 7.6$ 
\citep{Pustilnik06}; galaxies with high metallicities have $Z > 8.2$ 
\citep{Pilyugin06}.  The solar metallicity $Z_\odot = 8.86$ \citep{Delahaye06}.

@@

The selection criteria outlined in Section \ref{sec:SDSS_limits} limit our 
sample to only star-forming dwarf galaxies.  As a result, this is not a 
representative sample of the entire dwarf galaxy population.  We are only able 
to discuss the influence of the large-scale environment on star-forming dwarf 
galaxies in this study.  Unfortunately, it is impossible to estimate the 
chemical abundances of red dwarf galaxies with the direct $T_e$ method because 
the UV photons from young stars are needed to excite the interstellar gas.

@@

% Results table (machine-readable)
\floattable
\input{results_table_short}

The abundances estimated using the direct $T_e$ method for our dwarf galaxy 
sample are listed in Table \ref{tab:Results}.  Also included are other 
significant characteristics and identification for the galaxies, including their 
large-scale environmental classification.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.49\textwidth]{1sig_dwarf_SF_t3_12logOHrelations_dust_hist}
    \includegraphics[width=0.49\textwidth]{1sig_dwarf_SF_t3_12logOHrelations_dust_CDF}
    \caption{Gas-phase metallicity of void dwarf (red solid line) and wall 
    dwarf (black dashed line) galaxies.  A two-sample K-S test of the two data 
    sets results in an asymptotic $p$-value of 0.12, indicating a 12\% 
    probability that a test statistic greater than the observed value of 0.06 
    will be seen if the void sample is drawn from the wall sample.  This is 
    reflected visually, as there appears to be a slight but statistically 
    significant significant large-scale environmental influence on the 
    metallicity of dwarf galaxies.}
    \label{fig:met1sig}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=0.49\textwidth]{1sig_dwarf_SF_t3_12logNHrelations_dust_hist}
    \includegraphics[width=0.49\textwidth]{1sig_dwarf_SF_t3_12logNHrelations_dust_CDF}
    \caption{Abundance of nitrogen relative to hydrogen of void dwarf (red solid 
    line) and wall dwarf (black dashed line) galaxies.  A two-sample K-S test of 
    the two data sets results in an asymptotic $p$-value of 0.015, indicating a 
    1.5\% probability that a test statistic greater than the observed value of 
    0.08 will be seen, if the void sample is drawn from the wall sample.  This 
    is reflected visually, as the void dwarf galaxies appear to have lower 
    values of the N/H ratio than the wall dwarf galaxies.  There is a 
    large-scale environmental dependence of the chemical evolution of dwarf 
    galaxies.}
    \label{fig:N_1sig}
\end{figure*}

The distributions of oxygen and nitrogen abundances for dwarf galaxies as a 
function of large-scale environment are shown in Figures \ref{fig:met1sig} and 
\ref{fig:N_1sig}, respectively.  Both histograms show a slight shift between 
voids and walls in the chemical abundances of dwarf galaxies.  A two-sample 
Kolmogorov-Smirnov (K-S) test quantifies this observation --- a test statistic 
of 0.06 for oxygen and 0.08 for nitrogen are produced, corresponding to a 
probability of 12\% and 1.5\%, respectively, that a test statistic greater than 
or equal that observed will be measured if the void sample were drawn from the 
wall sample.  The cumulative distribution function (CDF) for each of these 
elements can be seen on the right in Figures \ref{fig:met1sig} and 
\ref{fig:N_1sig}; they show that void dwarf galaxies have slightly higher oxygen 
abundances and slightly lower nitrogen abundances than dwarf galaxies in more 
dense regions.  The K-S test quantifies the visual interpretation of these 
figures that the distributions of oxygen and nitrogen abundances are slightly 
different for star-forming dwarf galaxies in voids and walls.

The average and median values of the dwarf galaxy abundances also indicate a 
shift as a result of the large-scale environment.  The average oxygen abundance 
for void dwarf galaxies is $7.99\pm 0.007$ and the median is 8.06, while the 
average oxygen abundance for wall dwarf galaxies is $7.96\pm 0.009$ with a 
median value of 8.04.  This implies that the void dwarf galaxies have higher 
oxygen abundances by about 7\% (average shift of $0.03\pm 0.012$; median shift 
of 0.03) relative to wall dwarf galaxies.  In contrast, the average nitrogen 
abundance for void dwarf galaxies is only $6.55\pm 0.007$ with a median of 6.58, 
while the wall dwarf galaxies have an average nitrogen abundance of 
$6.58\pm 0.008$ and a median of 6.60.  The void dwarf galaxies have lower 
nitrogen abundances by about 10\% (an average shift of $0.04\pm 0.011$ and a 
median shift of 0.02) relative to wall dwarf galaxies.  A tabular version of 
this analysis can be found in Table \ref{tab:stats}.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.49\textwidth]{1sig_dwarf_SF_t3_logNOrelations_dust_hist}
    \includegraphics[width=0.49\textwidth]{1sig_dwarf_SF_t3_logNOrelations_dust_CDF}
    \caption{Ratio of nitrogen to oxygen of void dwarf (red solid line) and wall 
    dwarf (black dashed line) galaxies.  A two-sample K-S test of the two data 
    sets results in an asymptotic $p$-value of $2.6\times 10^{-4}$, indicating 
    only a 0.03\% probability that a test statistic greater than the observed 
    value of 0.10 will be seen.  This is reflected visually, as there is a shift 
    in the N/O ratio between the two populations of dwarf galaxies --- the void 
    galaxies have a lower value of N/O than the wall galaxies.  There is a 
    large-scale influence on the relative chemical abundances of dwarf 
    galaxies.}
    \label{fig:NOratio}
\end{figure*}

The ratio of nitrogen to oxygen is also important to investigate, as it 
communicates the nucleosynthesis history of the galaxies.  As seen in Fig. 
\ref{fig:NOratio}, the N/O abundance ratio also indicates a large-scale 
environmental influence on the chemical evolution of dwarf galaxies --- void 
dwarf galaxies have lower N/O ratios than dwarf galaxies in denser regions.  
This difference is quantified in the K-S test: the test returned a probability of 
only 0.03\% that a test statistic greater than or equal to 0.10 will be measured 
if the void sample was drawn from the wall sample.  The distribution of N/O 
abundance ratios for void dwarf galaxies is lower by about 17\% (an average 
shift of $0.07\pm 0.016$ and a median shift of 0.06) relative to the 
distribution of N/O ratios in wall dwarf galaxies.

% Statistics table
\floattable
\begin{deluxetable}{ccccccc}
    \tablewidth{0pt}
    \tablecolumns{7}
    \tablecaption{Abundance statistics\label{tab:stats}}
    \tablehead{\colhead{Environment} & \colhead{Average} & \colhead{Median} & \colhead{Average Shift\tablenotemark{a}} & \colhead{Median Shift\tablenotemark{a}} & \colhead{$p$-value} & \colhead{K-S Test Statistic}}
    \startdata
        % Oxygen %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \cutinhead{$12 + \log (\text{O}/\text{H})$}
        Void & $7.99\pm 0.007$ & 8.06 & \multirow{2}{*}{$-0.03\pm 0.012$} & \multirow{2}{*}{-0.03} & \multirow{2}{*}{0.1197} & \multirow{2}{*}{0.0569}\\
        Wall & $7.96\pm 0.009$ & 8.04 & & & & \\
        % Nitrogen %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \cutinhead{$12 + \log (\text{N}/\text{H})$}
        Void & $6.55\pm 0.007$ & 6.58 & \multirow{2}{*}{$0.04\pm 0.011$} & \multirow{2}{*}{0.02} & \multirow{2}{*}{0.0149} & \multirow{2}{*}{0.0750}\\
        Wall & $6.58\pm 0.008$ & 6.60 & & & & \\
        % N/O %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \cutinhead{$\log (\text{N}/\text{O})$}
        Void & $-1.45\pm 0.010$ & -1.43 & \multirow{2}{*}{$0.07\pm 0.016$} & \multirow{2}{*}{0.06} & \multirow{2}{*}{0.0003} & \multirow{2}{*}{0.1013}\\
        Wall & $-1.38\pm 0.012$ & -1.37 & & & & \\
    \enddata
    \tablecomments{Statistics on the gas-phase oxygen, nitrogen, and nitrogen 
    relative to oxygen abundances in dwarf void and wall galaxies.  Combined 
    with the histograms in Figures \ref{fig:met1sig}--\ref{fig:NOratio}, these 
    results indicate an influence on the chemical evolution of galaxies by the 
    large-scale environment, especially on the relative abundance of nitrogen to 
    oxygen.  Void galaxies have slightly higher oxygen and nitrogen abundances 
    than wall galaxies, but void galaxies have slightly lower N/O ratios than 
    wall galaxies.}
    \tablenotetext{a}{Wall -- Void (Positive shifts indicate that the wall 
    values are greater than the void values; negative shifts indicate that the 
    void values are greater than the wall values.)}
\end{deluxetable}

>>>>

We use the Direct $T_e$ method described in \cite{Izotov06} to estimate the 
gas-phase abundances of oxygen and nitrogen in our sample of dwarf galaxies, 
because this method is often regarded as the most accurate estimate of element 
abundances.  It can be difficult to use due to the nature of the [\ion{O}{3}] 
$\lambda$4363 auroral line \citep[for a more detailed discussion, see][]
{Douglass17a}.  Since the strength of [\ion{O}{3}] $\lambda$4363 is inversely 
proportional to the metallicity of a galaxy, it is best suited for low-redshift, 
low-metallicity galaxies.  At metallicities \OH $\gtrsim 8.5$, [\ion{O}{3}] 
$\lambda$4363 becomes too weak to detect in the SDSS spectra.  With the 
mass-metallicity (MZ) relation by \cite{Tremonti04}, this metallicity limit 
corresponds to a maximum stellar mass of
\begin{align*}
    12 + \log \left( \frac{\text{O}}{\text{H}} \right) &= -1.492 + 1.847(\log M_*) - 0.08026(\log M_*)^2\\
    8.5 &= \\
    \log \left( \frac{M_*}{M_\odot} \right) &= 8.696
\end{align*}
The MZ relation by \cite{Andrews13} estimates that this maximum metallicity 
translates to a maximum stellar mass of
\begin{align*}
    12 + \log \left( \frac{\text{O}}{\text{H}} \right) &= 12 + \log \left( \frac{\text{O}}{\text{H}} \right)_{asm} - \log \left( 1 + \left( \frac{M_{TO}}{M_*} \right)^\gamma \right)\\
    8.5 &= 8.798 - \log \left( 1 + \left( \frac{10^{8.901}}{M_*} \right)^{0.640} \right)\\
    M_* &= 8.138\times 10^8 M_\odot\\
    \log \left( \frac{M_*}{M_\odot} \right) &= 8.911
\end{align*}
These upper limits on the galaxy stellar masses correspond to the maximum mass 
of a dwarf galaxy in SDSS ($\log (M_*/M_\odot) \approx 9$).  Therefore, we can 
expect to estimate the chemical abundances of only dwarf galaxies in SDSS DR7 
with the Direct $T_e$ method.  Higher resolution spectra would be necessary to 
probe higher mass, higher metallicity galaxies.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.65\textwidth]{Images/Paper3/Zlow_v_Zreal_1sig_I06_BPTclass_fits}
    \caption[O$^++$/H versus O/H]{$12 + \log{\text{O}^{++}/\text{H}}$ versus \OH 
    for those galaxies with $M_r > -20$ as calculated with the direct method; 
    the galaxies are colored by their classification in the BPT diagrams 
    \citep[from][]{Brinchmann04}.  Two linear models have been fit to three 
    groups in the sample: star-forming and AGN/composite galaxies; the best fit 
    parameters can be found in Table \ref{tab:ab}.}
    \label{fig:lowVreal}
\end{figure}

We derive an ICF for O$^+$ to overcome some limitations on the observations of 
dwarf galaxies in SDSS to bolster our sample size.  Because we are studying 
dwarf galaxies, the [\ion{O}{2}] $\lambda 3727$ doublet necessary for estimating 
the abundance of O$^+$ is only available for objects within the redshift range 
$0.02 < z < 0.03$ \citep[see Sec. \ref{sec:SDSS_limits_P3} and][for more 
details]{Douglass17a}.  To be able to study SDSS galaxies at redshifts less than 
0.02, we need to find an alternate way to estimate the abundance of O$^+$.  
Ordinarily, we would use an ICF to replace the missing ionization stage, as 
outlined in Sec. \ref{sec:DirectTe} for the nitrogen abundance.  Because the 
[\ion{O}{2}] $\lambda 3727$ emission line is normally available for analysis, 
there are no approximations available for the ICF for the amount of 
singly-ionized oxygen.  To construct an ICF appropriate for our dwarf galaxy 
sample, we compare \OH to $12 + \log{\text{O}^{++}/\text{H}}$ as calculated by 
the Direct $T_e$ method for all galaxies in SDSS with $M_r > -20$.  Recognizing 
that the relative amounts of singly- and doubly-ionized oxygen will depend on 
the hardness of a galaxy's spectrum, we use a Baldwin-Phillips-Terlevich (BPT) 
diagram \citep{Baldwin81} to classify each galaxy as star-forming, AGN, or 
composite (containing characteristics of both a star-forming region and an AGN).  
Noting a strong separation in the relationship between O$^{++}$/H and O/H with 
respect to the galaxies' BPT classifications by \cite{Brinchmann04}, we fit two 
linear models to the sample, seen in Fig. \ref{fig:lowVreal} and defined as 
\begin{equation}\label{eq:fit}
    12 + \log{\text{O}/\text{H}} = a(12 + \log{\text{O}^{++}/\text{H}}) + b
\end{equation}
The values for $a$ and $b$ for the two classes are listed in Table \ref{tab:ab}.

\begin{table}
\centering

    \begin{tabular}{l|c|c|c}
         & SF (dwarf) & SF ($M_r > -20$) & AGN \& Composite\\
        \hline
        $a$ & $0.84 \pm 0.013$ & $0.824 \pm 0.0010$ & $0.764 \pm 0.0035$\\
        $b$ & $1.61 \pm 0.094$ & $1.787 \pm 0.0065$ & $2.05 \pm 0.023$
    \end{tabular}

    \caption[Coefficients of oxygen abundance fits]{Coefficients for the linear 
    trends (Eqn. \ref{eq:fit}) fit to the SF galaxies (dwarf and those with 
    $M_r > -20$) and AGN and composite galaxies.  With these trends, we can now 
    calculate the total oxygen abundance for a galaxy with knowing only the 
    O$^{++}$/H abundance.}

    \label{tab:ab}

\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%    DATA
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Data]{SDSS data and galaxy selection}

@@

%-------------------------------------------------------------------------------
\subsection{Spectroscopic selection}\label{sec:SDSS_limits_P3}

@@

%-------------------------------------------------------------------------------
\subsection{Void classification}

The large-scale environment of the galaxies is determined using the void catalog 
compiled by \cite{Pan12}, which was constructed with the galaxies in SDSS DR7 
catalog.  The VoidFinder algorithm of \cite{Hoyle02} \citep[based on the 
algorithm described by][]{ElAd97} removes all isolated galaxies (defined as 
having the third nearest neighbor more than 7 \hMpc away) using only galaxies 
with absolute magnitudes $M_r < -20$.  After applying a grid to the remaining 
galaxies, spheres are grown from all cells containing no galaxies until it 
encounters four galaxies on its surface.  A sphere must have a minimum 10 \hMpc 
radius to be classified as a void (or part of one).  If two spheres overlap by 
more than 10\%, they are considered part of the same void.  See \cite{Hoyle02} 
for a more detailed description of the VoidFinder algorithm.  Those galaxies 
that fall within these void spheres are classified as voids.  Galaxies that lie 
outside the spheres are classified as wall galaxies.  Because we cannot identify 
any voids within 5 \hMpc of the edge of the survey, we classify these galaxies 
as ``Uncertain.''

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%    ANALYSIS & RESULTS
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Results]{Abundance analysis and results}

@@

All line ratios listed are ratios of the emission-line fluxes.  Galaxies with 
low metallicities have $Z =$ \OH $< 7.6$ \citep{Pustilnik06}; galaxies with high 
metallicities have $Z > 8.2$ \citep{Pilyugin06}.  The solar metallicity 
$Z_\odot = 8.86$ \citep{Delahaye06}.

@@

The selection criteria outlined in Section \ref{sec:SDSS_limits_P3} limit our 
sample to only star-forming dwarf galaxies.  As a result, this is not a 
representative sample of the entire dwarf galaxy population.  We are only able 
to discuss the influence of the large-scale environment on star-forming dwarf 
galaxies in this study.  Unfortunately, it is impossible to estimate the 
chemical abundances of red dwarf galaxies with the direct $T_e$ method because 
the UV photons from young stars are needed to excite the interstellar gas.

@@

% Results table (machine-readable)
\input{Tables/Paper3/results_table_short}

The abundances estimated using the direct $T_e$ method for our dwarf galaxy 
sample are listed in Table \ref{tab:Results_P3}.  Also included are other 
significant characteristics and identification for the galaxies, including their 
large-scale environmental classification.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.49\textwidth]{Images/Paper3/1sig_dwarf_SF_t3_12logOHrelations_dust_hist}
    \includegraphics[width=0.49\textwidth]{Images/Paper3/1sig_dwarf_SF_t3_12logOHrelations_dust_CDF}
    \caption[O/H distribution for dwarf galaxy sample]{Gas-phase metallicity of 
    void dwarf (red solid line) and wall dwarf (black dashed line) galaxies.  A 
    two-sample K-S test of the two data sets results in an asymptotic $p$-value 
    of 0.12, indicating a 12\% probability that a test statistic greater than 
    the observed value of 0.06 will be seen if the void sample is drawn from the 
    wall sample.  This is reflected visually, as there appears to be a slight 
    but statistically significant significant large-scale environmental 
    influence on the metallicity of dwarf galaxies.}
    \label{fig:met1sig_P3}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=0.49\textwidth]{Images/Paper3/1sig_dwarf_SF_t3_12logNHrelations_dust_hist}
    \includegraphics[width=0.49\textwidth]{Images/Paper3/1sig_dwarf_SF_t3_12logNHrelations_dust_CDF}
    \caption[N/H distribution for dwarf galaxy sample]{Abundance of nitrogen 
    relative to hydrogen of void dwarf (red solid line) and wall dwarf (black 
    dashed line) galaxies.  A two-sample K-S test of the two data sets results 
    in an asymptotic $p$-value of 0.015, indicating a 1.5\% probability that a 
    test statistic greater than the observed value of 0.08 will be seen, if the 
    void sample is drawn from the wall sample.  This is reflected visually, as 
    the void dwarf galaxies appear to have lower values of the N/H ratio than 
    the wall dwarf galaxies.  There is a large-scale environmental dependence of 
    the chemical evolution of dwarf galaxies.}
    \label{fig:N_1sig_P3}
\end{figure*}

The distributions of oxygen and nitrogen abundances for dwarf galaxies as a 
function of large-scale environment are shown in Figures \ref{fig:met1sig_P3} 
and \ref{fig:N_1sig_P3}, respectively.  Both histograms show a slight shift 
between voids and walls in the chemical abundances of dwarf galaxies.  A 
two-sample Kolmogorov-Smirnov (K-S) test quantifies this observation --- a test 
statistic of 0.06 for oxygen and 0.08 for nitrogen are produced, corresponding 
to a probability of 12\% and 1.5\%, respectively, that a test statistic greater 
than or equal that observed will be measured if the void sample were drawn from 
the wall sample.  The cumulative distribution function (CDF) for each of these 
elements can be seen on the right in Figures \ref{fig:met1sig_P3} and 
\ref{fig:N_1sig_P3}; they show that void dwarf galaxies have slightly higher 
oxygen abundances and slightly lower nitrogen abundances than dwarf galaxies in 
more dense regions.  The K-S test quantifies the visual interpretation of these 
figures that the distributions of oxygen and nitrogen abundances are slightly 
different for star-forming dwarf galaxies in voids and walls.

The average and median values of the dwarf galaxy abundances also indicate a 
shift as a result of the large-scale environment.  The average oxygen abundance 
for void dwarf galaxies is $7.99\pm 0.007$ and the median is 8.06, while the 
average oxygen abundance for wall dwarf galaxies is $7.96\pm 0.009$ with a 
median value of 8.04.  This implies that the void dwarf galaxies have higher 
oxygen abundances by about 7\% (average shift of $0.03\pm 0.012$; median shift 
of 0.03) relative to wall dwarf galaxies.  In contrast, the average nitrogen 
abundance for void dwarf galaxies is only $6.55\pm 0.007$ with a median of 6.58, 
while the wall dwarf galaxies have an average nitrogen abundance of 
$6.58\pm 0.008$ and a median of 6.60.  The void dwarf galaxies have lower 
nitrogen abundances by about 10\% (an average shift of $0.04\pm 0.011$ and a 
median shift of 0.02) relative to wall dwarf galaxies.  A tabular version of 
this analysis can be found in Table \ref{tab:stats_p3}.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.49\textwidth]{Images/Paper3/1sig_dwarf_SF_t3_logNOrelations_dust_hist}
    \includegraphics[width=0.49\textwidth]{Images/Paper3/1sig_dwarf_SF_t3_logNOrelations_dust_CDF}
    \caption[N/O distribution of dwarf galaxy sample]{Ratio of nitrogen to 
    oxygen of void dwarf (red solid line) and wall dwarf (black dashed line) 
    galaxies.  A two-sample K-S test of the two data sets results in an 
    asymptotic $p$-value of $2.6\times 10^{-4}$, indicating only a 0.03\% 
    probability that a test statistic greater than the observed value of 0.10 
    will be seen.  This is reflected visually, as there is a shift in the N/O 
    ratio between the two populations of dwarf galaxies --- the void galaxies 
    have a lower value of N/O than the wall galaxies.  There is a large-scale 
    influence on the relative chemical abundances of dwarf galaxies.}
    \label{fig:NOratio_P3}
\end{figure*}

The ratio of nitrogen to oxygen is also important to investigate, as it 
communicates the nucleosynthesis history of the galaxies.  As seen in Fig. 
\ref{fig:NOratio_P3}, the N/O abundance ratio also indicates a large-scale 
environmental influence on the chemical evolution of dwarf galaxies --- void 
dwarf galaxies have lower N/O ratios than dwarf galaxies in denser regions.  
This difference is quantified in the K-S test: the test returned a probability of 
only 0.03\% that a test statistic greater than or equal to 0.10 will be measured 
if the void sample was drawn from the wall sample.  The distribution of N/O 
abundance ratios for void dwarf galaxies is lower by about 17\% (an average 
shift of $0.07\pm 0.016$ and a median shift of 0.06) relative to the 
distribution of N/O ratios in wall dwarf galaxies.

% Statistics table
\begin{table}
    \centering

    \begin{tabular}{ccccccc}
        Environment & Average & Median & Average Shift\footnote{Wall -- Void (Positive shifts indicate that the wall values are greater than the void values; negative shifts indicate that the void values are greater than the wall values.)} & Median Shift & $p$-value & K-S Test Statistic\\
        \hline
        % Oxygen %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \hline
        \multicolumn{7}{c}{\OH}\\
        \hline
        Void & $7.99\pm 0.007$ & 8.06 & \multirow{2}{*}{$-0.03\pm 0.012$} & \multirow{2}{*}{-0.03} & \multirow{2}{*}{0.1197} & \multirow{2}{*}{0.0569}\\
        Wall & $7.96\pm 0.009$ & 8.04 & & & & \\
        % Nitrogen %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \hline
        \multicolumn{7}{c}{\NH}\\
        \hline
        Void & $6.55\pm 0.007$ & 6.58 & \multirow{2}{*}{$0.04\pm 0.011$} & \multirow{2}{*}{0.02} & \multirow{2}{*}{0.0149} & \multirow{2}{*}{0.0750}\\
        Wall & $6.58\pm 0.008$ & 6.60 & & & & \\
        % N/O %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \hline
        \multicolumn{7}{c}{\NO}\\
        \hline
        Void & $-1.45\pm 0.010$ & -1.43 & \multirow{2}{*}{$0.07\pm 0.016$} & \multirow{2}{*}{0.06} & \multirow{2}{*}{0.0003} & \multirow{2}{*}{0.1013}\\
        Wall & $-1.38\pm 0.012$ & -1.37 & & & & \\
    \end{tabular}

    \caption[Abundance statistics]{Statistics on the gas-phase oxygen, nitrogen, 
    and nitrogen relative to oxygen abundances in dwarf void and wall galaxies.  
    Combined with the histograms in Figures 
    \ref{fig:met1sig_P3}--\ref{fig:NOratio_P3}, these results indicate an 
    influence on the chemical evolution of galaxies by the large-scale 
    environment, especially on the relative abundance of nitrogen to oxygen.  
    Void galaxies have slightly higher oxygen and nitrogen abundances than wall 
    galaxies, but void galaxies have slightly lower N/O ratios than wall 
    galaxies.}

    \label{tab:stats_p3}

\end{table}
